# Reproducible Research: Assignment 1

---

```{r setoptions, echo=FALSE, results="hide"}
library(knitr)
opts_chunk$set(echo=TRUE, results="asis")
```

## 1.Code for reading in the dataset and/or processing the data

Load the data.
```{r}
d<- read.csv (file = "activity.csv", header = TRUE)
d$date<-as.Date(d$date,format='%Y-%m-%d')

```

##2 Histogram of the total number of steps taken each day
```{r}
q2 <- aggregate(x = d[c("steps","interval")],
                     FUN = sum,
                     by = list(Group.date = d$date), na.rm=TRUE)

##  Histogram
hist(q2$steps,
     col = "lightgray",
     border = "black", 
     xlab = "Interval",
     main = "Histogram - Total Steps per day")
  abline(v = mean(q2$steps),
         col = "blue",
         lwd = 2)
  abline(v = median(q2$steps),
         col = "red",
         lwd = 2)
  legend(x = "topright", 
       c("Mean", "Median"),
       col = c("blue", "red"),
       lwd = c(2, 2, 2))
```

## 3. Mean and median number of steps taken each day
```{r}
q3_mean <- setNames(aggregate(x = d[c("steps")],
                FUN = mean,
                by = list(d$date), na.rm=TRUE), c("Date", "Mean"))

q3_median <- setNames(aggregate(x = d[c("steps")],
                    FUN = median,
                    by = list(d$date), na.rm=TRUE), c("Date", "Median"))

q3 <- merge(q3_mean,q3_median,by="Date")

print(q3)
```
## 4. Time series plot of the average number of steps taken
```{r}
q4_mean <- setNames(aggregate(x = d[c("steps")],
                             FUN = mean,
                             by = list(d$interval), na.rm=TRUE), c("Interval", "MeanStep"))

## Interval with Max Average Steps
q4_max<-subset(q4_mean, q4_mean$MeanStep==max(q4_mean$MeanStep, na.rm = TRUE))

## plot histogram
plot(q4_mean$Interval, q4_mean$MeanStep, type="l", main = "Steps per Interval", xlab= "Interval", ylab= "Steps(Mean)", col="red" ,lwd=1)
    abline(v = q4_max$Interval, col = "blue", lwd = 2)
    legend(x = "topright", 
           c("MaxSteps Interval"),
           col = c("blue"),
           lwd = c(2, 2, 2))
```
## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
q4_max<-subset(q4_mean, q4_mean$MeanStep==max(q4_mean$MeanStep, na.rm = TRUE))
    print(q4_max)
```

## 6. Code to describe and show a strategy for imputing missing data

```{r}
## original data to be stored in a new variable called 'q6'
q6 <- d

##NA values in column 'Steps' are replaced with global average
q6$steps[which(is.na(q6$steps))] <- mean(q6$steps, na.rm = TRUE)

## example
head(q6)
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
q7 <- aggregate(x = q6[c("steps")],
                FUN = sum,
                by = list(Group.date = d$date), na.rm=TRUE)

## stats
q7_mean <- setNames(aggregate(x = q7[c("steps")],
                             FUN = mean,
                             by = list( q7$Group.date), na.rm=TRUE), c("Date", "Mean"))

q7_median <- setNames(aggregate(x = q7[c("steps")],
                               FUN = median,
                               by = list(q7$Group.date), na.rm=TRUE), c("Date", "Median"))

q7_all <- merge(q7_mean,q7_median,by="Date")

hist(q7$steps,
       col = "lightgray",
       border = "black", 
       xlab = "SUM(steps)",
       main = "Histogram  Total Steps per day")
        abline(v = mean(q7$steps),
         col = "blue",
         lwd = 2)
        
        abline(v = median(q7$steps),
         col = "red",
         lty = 2,
         lwd = 2)
        
        legend(x = "topright", 
         c("Mean","Median"),
         col = c("blue","red"),
         lwd = c(2, 2, 2))
```
## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
q8<-q6
q8$day <- weekdays(q6$date)
q8$day <- ifelse(q8$day == "Saturday" | q8$day == "Sunday" ,"weekend", "weekday")
q8$day <- as.factor(q8$day)


## aggregated data for graphic
q8_sum <- setNames(aggregate(x = q8[c("steps")],
                             FUN = sum,
                             by = list(q8$interval, q8$day), na.rm=TRUE), c("Interval", "day", "SumStep"))

## subsetting data for graphics
q8_weekday<-subset(q8_sum, q8_sum$day=="weekday")
q8_weekend<-subset(q8_sum, q8_sum$day=="weekend")

## plot
  par(mfrow=c(2,1))
        plot(q8_weekday$Interval,q8_weekday$SumStep, main="Interval vs SumStep [weekdays]", type = "l", col = "blue", xlab = "Interval", ylab = "SumStep")
        plot(q8_weekend$Interval,q8_weekend$SumStep, main="Interval vs SumStep [weekends]", type = "l", col = "red", xlab = "Interval", ylab = "SumStep")
```
## 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report